Estimating individual subjective values of emotion regulation strategies

### 2022-06-20 06_Paper/ER-ED/Manuscript_ERED_Stage1.Rmd

---
title             : "Estimating individual subjective values of emotion regulation strategies"
shorttitle        : "Subjective values of emotion regulation strategies"
author:
  - name          : "Christoph Scheffel"
    orcid         : "0000-0001-5963-9229"
    affiliation   : "1"
    corresponding : yes
    address       : "Zellescher Weg 17, 01069 Dresden, Germany"
    email         : "christoph_scheffel@tu-dresden.de"
    equal_contrib : yes
    role          :         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Methodology
      - Funding acquisition
      - Formal analysis
      - Investigation
      - Project administration
      - Software
      - Visualization
      - Writing - original draft preparation
      - Writing - review & editing
  - name          : "Josephine Zerna"
    orcid         : "0000-0003-2892-884X"
    affiliation   : "1"
    equal_contrib : yes
    role          :
      - Conceptualization
      - Methodology
      - Funding acquisition
      - Investigation
      - Project administration
      - Software
      - Visualization
      - Writing - review & editing
  - name          : "Anne Gärtner"
    orcid         : "0000-0003-4296-963X"
    affiliation   : "1"
    role          :
      - Formal analysis
      - Writing - review & editing
  - name          : "Denise Dörfel"
    orcid         :
    affiliation   : "1"
    role          :
      - Conceptuatlization
      - Writing - review & editing
  - name          : "Alexander Strobel"
    orcid         : "0000-0002-9426-5397"
    affiliation   : "1"
    role          :
      - Conceptualization
      - Writing - review & editing 
affiliation       :
  - id            : "1"
    institution   : "Faculty of Psychology, Technische Universität Dresden, 01069 Dresden, Germany"
authornote        :
abstract          : |
 Individuals have a repertoire of emotion regulation (ER) strategies at their disposal, which they can use more or less flexibly. 
 In ER flexibility research, strategies that facilitate goal achievement are considered adaptive and therefore are subjectively valuable. 
 Individuals are motivated to reduce their emotional arousal effectively and to avoid cognitive effort. 
 Perceived costs of ER strategies in the form of effort, however, are highly subjective. 
 Subjective values (SVs) should therefore represent a trade-off between effectiveness and subjectively required cognitive effort. 
 However, SVs of ER strategies have not been determined so far. 
 We present a paradigm that is suitable for determining individual SVs of ER strategies. 
 Using a multilevel modelling approach, it will be investigated whether individual SVs can be explained by effectiveness (subjective arousal, facial muscle activity) and subjective effort. 
 Relations of SVs to personality traits will be explored.
keywords          : "emotion regulation, regulatory effort, effort discounting, registered report, specification curve analysis"
wordcount         : "6233"
bibliography      : ["ER_ED.bib"]
floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
csl               : nature.csl
header-includes   :
    - \usepackage{booktabs}
    - \usepackage{longtable}
    - \usepackage{array}
    - \usepackage{multirow}
    - \usepackage{wrapfig}
    - \usepackage{float}
    - \usepackage{colortbl}
    - \usepackage{pdflscape}
    - \usepackage{tabu}
    - \usepackage{threeparttable}
    - \usepackage{threeparttablex}
    - \usepackage[normalem]{ulem}
    - \usepackage{makecell}
    - \usepackage{xcolor}
    - \usepackage{setspace}\doublespacing
    - \usepackage[final]{pdfpages}
    - \usepackage{chngcntr}
editor_options: 
  markdown: 
    wrap: sentence
---

@@

\includepdf[pages={-}, landscape=true]{Supplement/StimList_Suppl.pdf}

>>>>

---
title             : "Estimating individual subjective values of emotion regulation strategies"
shorttitle        : "Subjective values of emotion regulation strategies"
author:
  - name          : "Christoph Scheffel"
    orcid         : "0000-0001-5963-9229"
    affiliation   : "1"
    corresponding : yes
    address       : "Zellescher Weg 17, 01069 Dresden, Germany"
    email         : "christoph_scheffel@tu-dresden.de"
    equal_contrib : yes
    role          :         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Methodology
      - Funding acquisition
      - Formal analysis
      - Investigation
      - Project administration
      - Software
      - Visualization
      - Writing - original draft preparation
      - Writing - review & editing
  - name          : "Josephine Zerna"
    orcid         : "0000-0003-2892-884X"
    affiliation   : "1"
    equal_contrib : yes
    role          :
      - Conceptualization
      - Methodology
      - Funding acquisition
      - Investigation
      - Project administration
      - Software
      - Visualization
      - Writing - review & editing
  - name          : "Anne Gärtner"
    orcid         : "0000-0003-4296-963X"
    affiliation   : "1"
    role          :
      - Formal analysis
      - Writing - review & editing
  - name          : "Denise Dörfel"
    orcid         :
    affiliation   : "1"
    role          :
      - Conceptuatlization
      - Writing - review & editing
  - name          : "Alexander Strobel"
    orcid         : "0000-0002-9426-5397"
    affiliation   : "1"
    role          :
      - Conceptualization
      - Writing - review & editing 
affiliation       :
  - id            : "1"
    institution   : "Faculty of Psychology, Technische Universität Dresden, 01069 Dresden, Germany"
authornote        :
abstract          : |
 Individuals have a repertoire of emotion regulation (ER) strategies at their disposal, which they can use more or less flexibly. 
 In ER flexibility research, strategies that facilitate goal achievement are considered adaptive and therefore are subjectively valuable. 
 Individuals are motivated to reduce their emotional arousal effectively and to avoid cognitive effort. 
 Perceived costs of ER strategies in the form of effort, however, are highly subjective. 
 Subjective values (SVs) should therefore represent a trade-off between effectiveness and subjectively required cognitive effort. 
 However, SVs of ER strategies have not been determined so far. 
 We present a paradigm that is suitable for determining individual SVs of ER strategies. 
 Using a multilevel modelling approach, it will be investigated whether individual SVs can be explained by effectiveness (subjective arousal, facial muscle activity) and subjective effort. 
 Relations of SVs to personality traits will be explored.
keywords          : "emotion regulation, regulatory effort, effort discounting, registered report, specification curve analysis"
wordcount         : "6289"
bibliography      : ["ER_ED.bib"]
floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
csl               : nature.csl
header-includes   :
    - \usepackage{booktabs}
    - \usepackage{longtable}
    - \usepackage{array}
    - \usepackage{multirow}
    - \usepackage{wrapfig}
    - \usepackage{float}
    - \usepackage{colortbl}
    - \usepackage{pdflscape}
    - \usepackage{tabu}
    - \usepackage{threeparttable}
    - \usepackage{threeparttablex}
    - \usepackage[normalem]{ulem}
    - \usepackage{makecell}
    - \usepackage{xcolor}
    - \usepackage{setspace}\doublespacing
    - \usepackage[final]{pdfpages}
    - \usepackage{chngcntr}
editor_options: 
  markdown: 
    wrap: sentence
---

@@

\includepdf[pages={-}, landscape=true]{Supplement/StimList_Suppl.pdf}
\counterwithin{figure}{section}
\counterwithin{table}{section}
\setcounter{section}{19}
\setcounter{figure}{0}
\setcounter{table}{1}

### 2022-06-20 06_Paper/ER-ED/Manuscript_ERED_Stage1.tex

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{emotion regulation, regulatory effort, effort discounting, registered report, specification curve analysis\newline\indent Word count: 6233}
\usepackage{lineno}

@@

The newly developed ER paradigm was tested in a pilot study with \(N=16\) participants (9 female; age: \(M = 24.1\text{ }\pm\text{ }SD = 3.6\)).
Regarding self-reported arousal, results showed significant higher subjective arousal for active viewing of negative compared to neutral pictures.
However, ER strategies did not lead to a reduction of subjective arousal compared to active viewing of negative pictures.
Regarding physiological responses, ER strategies were associated with reduced facial muscle activity of the \emph{corrugator} and \emph{levator} compared to active viewing of negative pictures.
In accordance with our previous study\textsuperscript{22}, we found that the use of ER strategies compared to active viewing was associated with increased subjective effort.
All results are detailed in the \protect\hyperlink{SupplementaryMaterial}{Supplementary Material}

@@

Pictures that will be used in the paradigm are selected from the Emotional Picture Set (EmoPicS)\textsuperscript{60} and the International Affective Picture System (IAPS)\textsuperscript{61}.
The 20 neutral pictures (Valence (V): \emph{M} \(\pm\) \emph{SD} = 4.81 \(\pm\) 0.51; Arousal (A): \emph{M} \(\pm\) \emph{SD} = 3 \(\pm\) 0.65) depicted content related to the categories persons, objects, and scenes.
Further, 100 negative pictures, featuring categories animals, body, disaster, disgust, injury, suffering, violence, and weapons, will be used.
An evolutionary algorithm\textsuperscript{62} is used to cluster these pictures into five sets with comparable valence and arousal values (set one: V: \emph{M} \(\pm\) \emph{SD} = 2.84 \(\pm\) 0.57, A: \emph{M} \(\pm\) \emph{SD} = 5.62 \(\pm\) 0.34; set two: V: \emph{M} \(\pm\) \emph{SD} = 2.64 \(\pm\) 0.46, A: \emph{M} \(\pm\) \emph{SD} = 5.58 \(\pm\) 0.35; set three: V: \emph{M} \(\pm\) \emph{SD} = 2.82 \(\pm\) 0.62, A: \emph{M} \(\pm\) \emph{SD} = 5.60 \(\pm\) 0.39; set four: V: \emph{M} \(\pm\) \emph{SD} = 2.65 \(\pm\) 0.75, A: \emph{M} \(\pm\) \emph{SD} = 5.61 \(\pm\) 0.41; set five: V: \emph{M} \(\pm\) \emph{SD} = 2.74 \(\pm\) 0.70, A: \emph{M} \(\pm\) \emph{SD} = 5.63 \(\pm\) 0.37).
A complete list of all pictures and their classification into sets can be found in supplementary material 1.
The five sets of negative pictures will be assigned randomly to the blocks.

@@

To explore the association between subjective effort (H5a), subjective arousal (H5b), subjective utility (H5c), and physiological responding (H5d,e) on SVs, a multilevel model (MLM) will be specified using the \(lmerTest\) package\textsuperscript{67}.
First, ER strategies will be recoded and centered for each subject according to their individual SVs: The strategy with the highest SV will be coded as -1, the strategy with the second highest SV 0, and the strategy with the lowest SV will be coded as 1.
Restricted maximum likelihood (REML) will be applied to fit the model.
A random slopes model of SVs including subjective effort (effort ratings), subjective arousal (arousal ratings), and physiological responses (\emph{corrugator} and \emph{levator} activity) as level-1-predictors will be specified.
\[
SV \sim strategy\ + \text{effort rating} + \text{arousal rating} + \text{utility} + corrugator \text{ activity} + levator \text{ activity} + (strategy|subject)
\]
Level-1-predictors will be centered within cluster\textsuperscript{68}.
Residuals of the final model will be inspected visually.
Intraclass correlation coefficient (ICC), \(\rho\), will be reported for each model (null model, as well as full model).

>>>>

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{emotion regulation, regulatory effort, effort discounting, registered report, specification curve analysis\newline\indent Word count: 6289}
\usepackage{lineno}

@@

The newly developed ER paradigm was tested in a pilot study with \(N=16\) participants (9 female; age: \(M = 24.1\text{ }\pm\text{ }SD = 3.6\)).
Regarding self-reported arousal, results showed significant higher subjective arousal for active viewing of negative compared to neutral pictures.
However, ER strategies did not lead to a reduction of subjective arousal compared to active viewing of negative pictures.
Regarding physiological responses, ER strategies were associated with reduced facial muscle activity of the \emph{corrugator} and \emph{levator} compared to active viewing of negative pictures.
In accordance with our previous study\textsuperscript{22}, we found that the use of ER strategies compared to active viewing was associated with increased subjective effort.
All results are detailed in the \protect\hyperlink{SupplementaryMaterial}{Supplementary Material}, figures S1 to S7 and table S2 to S8.

@@

Pictures that will be used in the paradigm are selected from the Emotional Picture Set (EmoPicS)\textsuperscript{60} and the International Affective Picture System (IAPS)\textsuperscript{61}.
The 20 neutral pictures (Valence (V): \emph{M} \(\pm\) \emph{SD} = 4.81 \(\pm\) 0.51; Arousal (A): \emph{M} \(\pm\) \emph{SD} = 3 \(\pm\) 0.65) depicted content related to the categories persons, objects, and scenes.
Further, 100 negative pictures, featuring categories animals, body, disaster, disgust, injury, suffering, violence, and weapons, will be used.
An evolutionary algorithm\textsuperscript{62} is used to cluster these pictures into five sets with comparable valence and arousal values (set one: V: \emph{M} \(\pm\) \emph{SD} = 2.84 \(\pm\) 0.57, A: \emph{M} \(\pm\) \emph{SD} = 5.62 \(\pm\) 0.34; set two: V: \emph{M} \(\pm\) \emph{SD} = 2.64 \(\pm\) 0.46, A: \emph{M} \(\pm\) \emph{SD} = 5.58 \(\pm\) 0.35; set three: V: \emph{M} \(\pm\) \emph{SD} = 2.82 \(\pm\) 0.62, A: \emph{M} \(\pm\) \emph{SD} = 5.60 \(\pm\) 0.39; set four: V: \emph{M} \(\pm\) \emph{SD} = 2.65 \(\pm\) 0.75, A: \emph{M} \(\pm\) \emph{SD} = 5.61 \(\pm\) 0.41; set five: V: \emph{M} \(\pm\) \emph{SD} = 2.74 \(\pm\) 0.70, A: \emph{M} \(\pm\) \emph{SD} = 5.63 \(\pm\) 0.37).
A complete list of all pictures and their classification into sets can be found in supplementary material table S1.
The five sets of negative pictures will be assigned randomly to the blocks.

@@

To explore the association between subjective effort (H5a), subjective arousal (H5b), subjective utility (H5c), and physiological responding (H5d,e) on SVs, a multilevel model (MLM) will be specified using the \(lmerTest\) package\textsuperscript{67}.
First, ER strategies will be recoded and centered for each subject according to their individual SVs: The strategy with the highest SV will be coded as -1, the strategy with the second highest SV 0, and the strategy with the lowest SV will be coded as 1.
Restricted maximum likelihood (REML) will be applied to fit the model.
A random slopes model of SVs including subjective effort (effort ratings), subjective arousal (arousal ratings), utility (utility ratings), and physiological responses (\emph{corrugator} and \emph{levator} activity) as level-1-predictors will be specified.
\[
\begin{split}
SV \sim strategy\ + \text{effort rating} + \text{arousal rating} + \text{utility rating} + corrugator \text{ activity} \\\ + levator \text{ activity} + (strategy|subject)
\end{split}
\]
Level-1-predictors will be centered within cluster\textsuperscript{68}.
Residuals of the final model will be inspected visually.
Intraclass correlation coefficient (ICC), \(\rho\), will be reported for each model (null model, as well as full model).
